Imputation of Incomplete Multilevel Data with R

### 2023-12-12 docs/manuscript.qmd
Case study 3

```{R}
#| prompt: true
#| warning: false
#| message: false
set.seed(123)         # for reproducibility
library(lme4)         # for multilevel modeling
library(mice)         # for imputation
library(miceadds)     # for multilevel imputation methods
library(micemd)       # for selection-model imputation methods
library(mitml)        # for multilevel parameter pooling
library(broom)        # for clean model estimates
library(broom.mixed)  # for multilevel model estimates
library(dplyr)        # for data wrangling
library(ggmice)       # for visualization
library(ggplot2)      # for visualization
```

@@

```{r}
#| prompt: true
Obesity |> 
  split(~Cluster) |>
  lapply(plot_pattern)
```

@@

```{r obesity-icc, eval=TRUE}
# Nulmodel <- lme4::lmer(BMI ~ 1 + (1|Cluster), data = Obesity)
# performance::icc(Nulmodel)
```

@@

```{r obesitymeth, eval=TRUE}
# meth_mar <- micemd::find.defaultMethod(Obesity, ind.clust=1, I.small = 7,
#                                ni.small = 100, prop.small = 0.4)
# meth_mar["BMI"]<- "~ I(Weight / (Height)^2)"
# meth_mar["Age"]<-"2l.2stage.pmm" 
```

For these imputation models, it is necessary to specify the prediction matrix, with the cluster
variable labelled as -2 and the predictor variable measured within clusters labelled as 2, encompassing all variables. We need to suprime the variable Time as this variable is not specified in the main model. We also  modify the relationship between BMI, weight and height in the prediction matrix to avoid circular predictions. Then we proceed to run the imputation model.
```{r obesity-predmar,out.width="70%", eval=TRUE}
# pred_mar <- mice(Obesity, maxit = 0)$pred 
# pred_mar[,"Cluster"] <- -2 # clustering variable
# pred_mar[,"Time"] <- 0
# pred_mar[pred_mar==1] <- 2
# pred_mar[c("Height", "Weight"), "BMI"] <- 0
# ggmice::plot_pred(pred_mar)
# imp_mar <- mice::mice(data = Obesity, meth = meth_mar, pred = pred_mar,
#                 m=10, seed = 123, printFlag = FALSE)
```

```{r obesity-predmar1,eval=TRUE}
# summary(complete(imp_mar,"long")$Weight)
```
We are also contemplating the utilisation of the predictive mean matching (pmm) option, as the values imputed using a fully parametric method may be implausibly low for some patients.

```{r obesity-predmar_pmm, eval=TRUE}
# meth_mar["Weight"]<-"2l.2stage.pmm" 
# imp_mar_pmm <- mice(data = Obesity, meth = meth_mar, pred = pred_mar,
#                     m=10, seed = 123, printFlag = FALSE)
```

```{r obesity-predmar_pmm1,eval=TRUE}
# summary(complete(imp_mar_pmm,"long")$Weight)
# ggmice::plot_trace(imp_mar_pmm, "Weight")
```

@@

```{r obesity-predmnar, eval=TRUE}
# pred_mnar <- pred_mar
# pred_mnar["Weight","Time"]<- -3
# ggmice::plot_pred(pred_mnar)
```

@@

```{r obesity-predmnar1, eval=TRUE}
# meth_mnar <- meth_mar
# meth_mnar["Weight"]<- "3l.2stage.heckman"
```
Then we proceed to run the imputation model as before, after executing these imputation procedures, it is essential to assess convergence and the coherence of the imputed values.
```{r obesity-predmnar2,eval=TRUE}
# imp_mnar<- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
#                 m=10, seed = 123, printFlag = FALSE)
# summary(complete(imp_mnar,"long")$Weight)
```

@@

```{r obesity-predmnar3,eval=TRUE}
# summary(complete(imp_mnar,"long")$Weight)
```
Consequently, as before we  use the 'pmm', option but this time for the Heckman imputation, this approach ensures that the imputed values remain within the range of observable values. We then run the imputation model but this time using the option of pmm, to assure that weight values are in the range of the observable data, this can be implemented by setting the pmm parameter to true.

```{r obesity-predmnarp, eval=TRUE}
# imp_mnar_pmm <- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
#                      m=10, seed = 123, pmm = T,  printFlag = FALSE)
```

We check the convergency of the results
```{r obesity-predmnarp1, eval=TRUE}
# summary(complete(imp_mnar_pmm,"long")$Weight)
# ggmice::plot_trace(imp_mar_pmm, "Weight")
```

After this modification we proceed to compare the effects on the model. We
run the analysis model on each of the completed datasets as well asthe dataset where the
incomplete values are removed (Complete Case analysis, CC).

```{r models, eval=TRUE}
# cc_rs<- with(setDT(Obesity)[complete.cases(Obesity),],
#              lme( BMI ~ Age + FamOb + Gender, random=~1+Age|Cluster))
# mar_rs <- with(imp_mar,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# mar_pmm_rs <- with(imp_mar_pmm,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# mnar_rs<- with(imp_mnar,lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# mnar_pmm_rs<- with(imp_mnar_pmm, lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# list_models<-list(cc_rs,mar_rs,mar_pmm_rs,mnar_rs,mnar_pmm_rs)
# plot_models(list_models,
#             mod_name = c("Complete case", "MAR","MAR_pmm", "MNAR", "MNAR_pmm"))
```

We note that there is minimal disparity in the age effect, FamObs, or  Gender across the various imputation models under consideration. An analysis of the intercept reveals that, under the MNAR assumption, a higher average BMI is anticipated compared to the MAR assumption.
Nonetheless, with respect to precision of estimates, we notice that in general MNAR imputation leads to wider confidence intervals, in this case it does not have any influence on the final result but there could be cases where variation in the assumed missing mechanism could lead also to differences on significant test and therefore lead to contradictory conclusions.

>>>>

```{R}
#| prompt: true
#| warning: false
#| message: false
set.seed(123)         # for reproducibility
library(lme4)         # for multilevel modeling
library(nlme)         # for multilevel modeling
library(mice)         # for imputation
library(miceadds)     # for multilevel imputation methods
library(micemd)       # for selection-model imputation methods
library(mitml)        # for multilevel parameter pooling
library(broom)        # for clean model estimates
library(broom.mixed)  # for multilevel model estimates
library(dplyr)        # for data wrangling
library(ggmice)       # for visualization
library(ggplot2)      # for visualization
```

@@

```{r}
#| prompt: true
# Obesity |> 
#   split(~Cluster) |>
#   lapply(plot_pattern)
```

@@

```{r}
#| prompt: true
Nulmodel <- lmer(BMI ~ 1 + (1|Cluster), data = Obesity)
# performance::icc(Nulmodel)
```

@@

```{r obesitymeth, eval=TRUE}
meth_mar <-
  find.defaultMethod(
    Obesity,
    ind.clust = 1,
    I.small = 7,
    ni.small = 100,
    prop.small = 0.4
  )
meth_mar["BMI"] <- "~ I(Weight / (Height)^2)"
meth_mar["Age"] <- "2l.2stage.pmm"
```

For these imputation models, it is necessary to specify the prediction matrix, with the cluster
variable labelled as -2 and the predictor variable measured within clusters labelled as 2, encompassing all variables. We need to suprime the variable Time as this variable is not specified in the main model. We also  modify the relationship between BMI, weight and height in the prediction matrix to avoid circular predictions. Then we proceed to run the imputation model.
```{r obesity-predmar,out.width="70%", eval=TRUE}
pred_mar <- quickpred(Obesity)
pred_mar[, "Cluster"] <- -2 
pred_mar[, "Time"] <- 0
pred_mar[pred_mar == 1] <- 2
pred_mar[c("Height", "Weight"), "BMI"] <- 0
plot_pred(pred_mar)
imp_mar <-
  mice(
    data = Obesity,
    method = meth_mar,
    predictorMatrix = pred_mar,
    printFlag = FALSE
  )
```

```{r obesity-predmar1,eval=TRUE}
summary(complete(imp_mar,"long")$Weight)
```
We are also contemplating the utilisation of the predictive mean matching (pmm) option, as the values imputed using a fully parametric method may be implausibly low for some patients.

```{r obesity-predmar_pmm, eval=TRUE}
meth_mar["Weight"] <- "2l.2stage.pmm"
imp_mar_pmm <- mice(
  data = Obesity,
  method = meth_mar,
  predictorMatrix = pred_mar,
  printFlag = FALSE
)
```

```{r obesity-predmar_pmm1,eval=TRUE}
summary(complete(imp_mar_pmm, "long")$Weight)
plot_trace(imp_mar_pmm, "Weight")
ggmice(imp_mar_pmm, aes(.imp, Weight)) + 
  geom_jitter()
```

Analyze the imputed data.
```{R}
#| prompt: true
#| warning: false
# fit <- with(
#     imp_mar_pmm,
#     lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
#     )
fit <- with(
    imp_mar_pmm,
    lmer(BMI ~ Age + FamOb + Gender + (1 + Age | Cluster))
    )
est <- testEstimates(as.mitml.result(fit), extra.pars = TRUE)
```

@@

```{r obesity-predmnar, eval=TRUE}
pred_mnar <- pred_mar
pred_mnar["Weight","Time"] <- -3
plot_pred(pred_mnar)
```

@@

```{r obesity-predmnar1, eval=TRUE}
meth_mnar <- meth_mar
meth_mnar["Weight"] <- "2l.2stage.heckman"
```
Then we proceed to run the imputation model as before, after executing these imputation procedures, it is essential to assess convergence and the coherence of the imputed values.
```{r obesity-predmnar2,eval=TRUE}
imp_mnar <- mice(
  data = Obesity,
  method = meth_mnar,
  predictorMatrix = pred_mnar,
  printFlag = FALSE
)
summary(complete(imp_mnar, "long")$Weight)
```

@@

```{r obesity-predmnar3,eval=TRUE}
summary(complete(imp_mnar, "long")$Weight)
```
Consequently, as before we  use the 'pmm', option but this time for the Heckman imputation, this approach ensures that the imputed values remain within the range of observable values. We then run the imputation model but this time using the option of pmm, to assure that weight values are in the range of the observable data, this can be implemented by setting the pmm parameter to true.

```{r obesity-predmnarp, eval=TRUE}
imp_mnar_pmm <-
  mice(
    data = Obesity,
    method = meth_mnar,
    predictorMatrix = pred_mnar,
    pmm = T,
    printFlag = FALSE
  )
```

We check the convergence of the results
```{r obesity-predmnarp1, eval=TRUE}
summary(complete(imp_mnar_pmm, "long")$Weight)
plot_trace(imp_mar_pmm, "Weight")
```

After this modification we proceed to compare the effects on the model. We
run the analysis model on each of the completed datasets as well as the dataset where the
incomplete values are removed (Complete Case analysis, CC).

```{r models, eval=TRUE}
# cc_rs <- with(setDT(Obesity)[complete.cases(Obesity), ],
#               lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age |
#                     Cluster))
mar_rs <- with(
  imp_mar, 
  lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
  )
mar_pmm_rs <- with(
    imp_mar_pmm,
    lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
    )
mnar_rs <- with(
  imp_mnar, 
  lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
  )
mnar_pmm_rs <- with(
  imp_mnar_pmm,
  lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
  )
# list_models <- list(cc_rs, mar_rs, mar_pmm_rs, mnar_rs, mnar_pmm_rs)
# plot_models(list_models,
#             mod_name = c("Complete case", "MAR", "MAR_pmm", "MNAR", "MNAR_pmm"))
```

```{r}
est <- testEstimates(as.mitml.result(mnar_pmm_rs), extra.pars = TRUE)
```

```{r}
#| prompt: true
est$extra.pars |> 
  round(3) |>
  kable()
```
We note that there is minimal disparity in the age effect, FamObs, or  Gender across the various imputation models under consideration. An analysis of the intercept reveals that, under the MNAR assumption, a higher average BMI is anticipated compared to the MAR assumption.
Nonetheless, with respect to precision of estimates, we notice that in general MNAR imputation leads to wider confidence intervals, in this case it does not have any influence on the final result but there could be cases where variation in the assumed missing mechanism could lead also to differences on significant test and therefore lead to contradictory conclusions.

### 2023-12-12 docs/manuscript.tex
Case study 3

\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, breakable, toprule=.15mm, rightrule=.15mm, arc=.35mm, left=2mm, leftrule=.75mm, colback=white, opacityback=0]

@@

\begin{verbatim}
R> set.seed(123)         # for reproducibility
R> library(lme4)         # for multilevel modeling
R> library(mice)         # for imputation
R> library(miceadds)     # for multilevel imputation methods
R> library(micemd)       # for selection-model imputation methods
R> library(mitml)        # for multilevel parameter pooling
R> library(broom)        # for clean model estimates
R> library(broom.mixed)  # for multilevel model estimates
R> library(dplyr)        # for data wrangling
R> library(ggmice)       # for visualization
R> library(ggplot2)      # for visualization
\end{verbatim}

@@

\begin{verbatim}
R> Obesity |> 
+  split(~Cluster) |>
+  lapply(plot_pattern)
\end{verbatim}

\begin{verbatim}
$`1`
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/unnamed-chunk-49-1.pdf}

\begin{verbatim}

$`2`
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/unnamed-chunk-49-2.pdf}

\begin{verbatim}

$`3`
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/unnamed-chunk-49-3.pdf}

\begin{verbatim}

$`4`
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/unnamed-chunk-49-4.pdf}

\begin{verbatim}

$`5`
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/unnamed-chunk-49-5.pdf}

@@

\begin{verbatim}
# Nulmodel <- lme4::lmer(BMI ~ 1 + (1|Cluster), data = Obesity)
# performance::icc(Nulmodel)
\end{verbatim}

@@

\begin{verbatim}
# meth_mar <- micemd::find.defaultMethod(Obesity, ind.clust=1, I.small = 7,
#                                ni.small = 100, prop.small = 0.4)
# meth_mar["BMI"]<- "~ I(Weight / (Height)^2)"
# meth_mar["Age"]<-"2l.2stage.pmm" 
\end{verbatim}

@@

\begin{verbatim}
# pred_mar <- mice(Obesity, maxit = 0)$pred 
# pred_mar[,"Cluster"] <- -2 # clustering variable
# pred_mar[,"Time"] <- 0
# pred_mar[pred_mar==1] <- 2
# pred_mar[c("Height", "Weight"), "BMI"] <- 0
# ggmice::plot_pred(pred_mar)
# imp_mar <- mice::mice(data = Obesity, meth = meth_mar, pred = pred_mar,
#                 m=10, seed = 123, printFlag = FALSE)
\end{verbatim}

\begin{verbatim}
# summary(complete(imp_mar,"long")$Weight)
\end{verbatim}

@@

\begin{verbatim}
# meth_mar["Weight"]<-"2l.2stage.pmm" 
# imp_mar_pmm <- mice(data = Obesity, meth = meth_mar, pred = pred_mar,
#                     m=10, seed = 123, printFlag = FALSE)
\end{verbatim}

\begin{verbatim}
# summary(complete(imp_mar_pmm,"long")$Weight)
# ggmice::plot_trace(imp_mar_pmm, "Weight")
\end{verbatim}

@@

\begin{verbatim}
# pred_mnar <- pred_mar
# pred_mnar["Weight","Time"]<- -3
# ggmice::plot_pred(pred_mnar)
\end{verbatim}

@@

\begin{verbatim}
# meth_mnar <- meth_mar
# meth_mnar["Weight"]<- "3l.2stage.heckman"
\end{verbatim}

@@

\begin{verbatim}
# imp_mnar<- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
#                 m=10, seed = 123, printFlag = FALSE)
# summary(complete(imp_mnar,"long")$Weight)
\end{verbatim}

@@

\begin{verbatim}
# summary(complete(imp_mnar,"long")$Weight)
\end{verbatim}

@@

\begin{verbatim}
# imp_mnar_pmm <- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
#                      m=10, seed = 123, pmm = T,  printFlag = FALSE)
\end{verbatim}

We check the convergency of the results

\begin{verbatim}
# summary(complete(imp_mnar_pmm,"long")$Weight)
# ggmice::plot_trace(imp_mar_pmm, "Weight")
\end{verbatim}

After this modification we proceed to compare the effects on the model.
We run the analysis model on each of the completed datasets as well
asthe dataset where the incomplete values are removed (Complete Case
analysis, CC).

\begin{verbatim}
# cc_rs<- with(setDT(Obesity)[complete.cases(Obesity),],
#              lme( BMI ~ Age + FamOb + Gender, random=~1+Age|Cluster))
# mar_rs <- with(imp_mar,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# mar_pmm_rs <- with(imp_mar_pmm,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# mnar_rs<- with(imp_mnar,lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# mnar_pmm_rs<- with(imp_mnar_pmm, lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
# list_models<-list(cc_rs,mar_rs,mar_pmm_rs,mnar_rs,mnar_pmm_rs)
# plot_models(list_models,
#             mod_name = c("Complete case", "MAR","MAR_pmm", "MNAR", "MNAR_pmm"))
\end{verbatim}

>>>>

\begin{tcolorbox}[enhanced jigsaw, breakable, colback=white, opacityback=0, left=2mm, toprule=.15mm, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, leftrule=.75mm]

@@

\begin{verbatim}
R> set.seed(123)         # for reproducibility
R> library(lme4)         # for multilevel modeling
R> library(nlme)         # for multilevel modeling
R> library(mice)         # for imputation
R> library(miceadds)     # for multilevel imputation methods
R> library(micemd)       # for selection-model imputation methods
R> library(mitml)        # for multilevel parameter pooling
R> library(broom)        # for clean model estimates
R> library(broom.mixed)  # for multilevel model estimates
R> library(dplyr)        # for data wrangling
R> library(ggmice)       # for visualization
R> library(ggplot2)      # for visualization
\end{verbatim}

@@

\begin{verbatim}
R> # Obesity |> 
R> #   split(~Cluster) |>
R> #   lapply(plot_pattern)
\end{verbatim}

@@

\begin{verbatim}
R> Nulmodel <- lmer(BMI ~ 1 + (1|Cluster), data = Obesity)
R> # performance::icc(Nulmodel)
\end{verbatim}

@@

\begin{verbatim}
meth_mar <-
  find.defaultMethod(
    Obesity,
    ind.clust = 1,
    I.small = 7,
    ni.small = 100,
    prop.small = 0.4
  )
meth_mar["BMI"] <- "~ I(Weight / (Height)^2)"
meth_mar["Age"] <- "2l.2stage.pmm"
\end{verbatim}

@@

\begin{verbatim}
pred_mar <- quickpred(Obesity)
pred_mar[, "Cluster"] <- -2 
pred_mar[, "Time"] <- 0
pred_mar[pred_mar == 1] <- 2
pred_mar[c("Height", "Weight"), "BMI"] <- 0
plot_pred(pred_mar)
\end{verbatim}

{\centering \includegraphics[width=0.7\textwidth,height=\textheight]{manuscript_files/figure-pdf/obesity-predmar-1.pdf}

\begin{verbatim}
imp_mar <-
  mice(
    data = Obesity,
    method = meth_mar,
    predictorMatrix = pred_mar,
    printFlag = FALSE
  )
\end{verbatim}

\begin{verbatim}
summary(complete(imp_mar,"long")$Weight)
\end{verbatim}

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  4.428  69.890  82.800  82.835  95.491 145.885 
\end{verbatim}

@@

\begin{verbatim}
meth_mar["Weight"] <- "2l.2stage.pmm"
imp_mar_pmm <- mice(
  data = Obesity,
  method = meth_mar,
  predictorMatrix = pred_mar,
  printFlag = FALSE
)
\end{verbatim}

\begin{verbatim}
summary(complete(imp_mar_pmm, "long")$Weight)
\end{verbatim}

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  28.35   71.04   83.65   83.87   95.66  134.61 
\end{verbatim}

\begin{verbatim}
plot_trace(imp_mar_pmm, "Weight")
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/obesity-predmar_pmm1-1.pdf}

\begin{verbatim}
ggmice(imp_mar_pmm, aes(.imp, Weight)) + 
  geom_jitter()
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/obesity-predmar_pmm1-2.pdf}

\begin{verbatim}
R> # fit <- with(
R> #     imp_mar_pmm,
R> #     lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
R> #     )
R> fit <- with(
+    imp_mar_pmm,
+    lmer(BMI ~ Age + FamOb + Gender + (1 + Age | Cluster))
+    )
R> est <- testEstimates(as.mitml.result(fit), extra.pars = TRUE)
\end{verbatim}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1558}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1169}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1299}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1039}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1039}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.2338}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0779}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0779}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Std.Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
t.value
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
df
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
P(\textgreater\textbar t\textbar)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
RIV
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FMI
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(Intercept) & 28.445 & 2.262 & 12.578 & 158.191 & 0.000 & 0.189 &
0.169 \\
Age & 0.041 & 0.032 & 1.312 & 383.212 & 0.190 & 0.114 & 0.107 \\
FamObyes & 0.000 & 0.316 & -0.001 & 207.642 & 0.999 & 0.161 & 0.147 \\
GenderMale & -1.361 & 0.411 & -3.310 & 7.985 & 0.011 & 2.422 & 0.761 \\
\end{longtable}

\begin{longtable}[]{@{}lr@{}}
\toprule\noalign{}
& Estimate \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Intercept\textasciitilde\textasciitilde Intercept\textbar Cluster &
19.862 \\
Age\textasciitilde\textasciitilde Age\textbar Cluster & 0.002 \\
Intercept\textasciitilde\textasciitilde Age\textbar Cluster & -0.067 \\
Residual\textasciitilde\textasciitilde Residual & 26.010 \\
ICC\textbar Cluster & 0.395 \\
\end{longtable}

@@

\begin{verbatim}
pred_mnar <- pred_mar
pred_mnar["Weight","Time"] <- -3
plot_pred(pred_mnar)
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/obesity-predmnar-1.pdf}

@@

\begin{verbatim}
meth_mnar <- meth_mar
meth_mnar["Weight"] <- "2l.2stage.heckman"
\end{verbatim}

@@

\begin{verbatim}
imp_mnar <- mice(
  data = Obesity,
  method = meth_mnar,
  predictorMatrix = pred_mnar,
  printFlag = FALSE
)
summary(complete(imp_mnar, "long")$Weight)
\end{verbatim}

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  16.26   70.95   85.21   86.14  100.14  190.95 
\end{verbatim}

@@

\begin{verbatim}
summary(complete(imp_mnar, "long")$Weight)
\end{verbatim}

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  16.26   70.95   85.21   86.14  100.14  190.95 
\end{verbatim}

@@

\begin{verbatim}
imp_mnar_pmm <-
  mice(
    data = Obesity,
    method = meth_mnar,
    predictorMatrix = pred_mnar,
    pmm = T,
    printFlag = FALSE
  )
\end{verbatim}

We check the convergence of the results

\begin{verbatim}
summary(complete(imp_mnar_pmm, "long")$Weight)
\end{verbatim}

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  28.35   71.51   85.07   85.35   98.90  134.61 
\end{verbatim}

\begin{verbatim}
plot_trace(imp_mar_pmm, "Weight")
\end{verbatim}

{\centering \includegraphics{manuscript_files/figure-pdf/obesity-predmnarp1-1.pdf}

After this modification we proceed to compare the effects on the model.
We run the analysis model on each of the completed datasets as well as
the dataset where the incomplete values are removed (Complete Case
analysis, CC).

\begin{verbatim}
# cc_rs <- with(setDT(Obesity)[complete.cases(Obesity), ],
#               lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age |
#                     Cluster))
mar_rs <- with(
  imp_mar, 
  lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
  )
mar_pmm_rs <- with(
    imp_mar_pmm,
    lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
    )
mnar_rs <- with(
  imp_mnar, 
  lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
  )
mnar_pmm_rs <- with(
  imp_mnar_pmm,
  lme(BMI ~ Age + FamOb + Gender, random =  ~ 1 + Age | Cluster)
  )
# list_models <- list(cc_rs, mar_rs, mar_pmm_rs, mnar_rs, mnar_pmm_rs)
# plot_models(list_models,
#             mod_name = c("Complete case", "MAR", "MAR_pmm", "MNAR", "MNAR_pmm"))
\end{verbatim}

\begin{verbatim}
est <- testEstimates(as.mitml.result(mnar_pmm_rs), extra.pars = TRUE)
\end{verbatim}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1558}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1169}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1299}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1039}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0909}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.2338}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0909}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0779}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Std.Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
t.value
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
df
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
P(\textgreater\textbar t\textbar)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
RIV
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
FMI
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(Intercept) & 28.880 & 2.391 & 12.078 & 56.876 & 0.000 & 0.361 &
0.290 \\
Age & 0.050 & 0.037 & 1.343 & 19.874 & 0.194 & 0.814 & 0.497 \\
FamObyes & -0.027 & 0.490 & -0.056 & 11.685 & 0.957 & 1.410 & 0.642 \\
GenderMale & -1.484 & 0.804 & -1.846 & 4.813 & 0.126 & 10.318 & 0.934 \\
\end{longtable}

\begin{longtable}[]{@{}lr@{}}
\toprule\noalign{}
& Estimate \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Intercept\textasciitilde\textasciitilde Intercept\textbar Cluster &
19.111 \\
Age\textasciitilde\textasciitilde Age\textbar Cluster & 0.001 \\
Intercept\textasciitilde\textasciitilde Age\textbar Cluster & -0.046 \\
Residual\textasciitilde\textasciitilde Residual & 30.071 \\
ICC\textbar Cluster & 0.372 \\
\end{longtable}
